Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-17 paper/paper.tex
Christine's changes

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs~\cite{constraints10}.
\item Classes 2 and 3 contain instances built from a database containing
    various kinds of graph gathered by Larrosa and Valiente~\cite{LV02}: class
    2 contains small instances generated from the first 50 graphs of the
    database, and class 3 contains larger instances with pattern graphs from
    the first 50 graphs of the database and target graphs from the next 50
    graphs.
\item Classes 4 to 8 contain randomly generated graphs from a database of
    graphs commonly used for benchmarking subgraph isomorphism
    algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for
    classes 4 and 5, regular meshes for classes 6 and 7, and random graphs with
    uniform edge probabilities for class 8. All of these instances are
    satisfiable.
\item Classes 9 and 10 contain instances from segmented images~\cite{pr15,cviu11}.
\item Class 11 contains instances from meshes modeling 3D objects~\cite{cviu11}.
\item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
    phase transition---these instances are expected to be particularly challenging, despite their
    small size.
\end{itemize}

\noindent Note that Classes 3 and 12 were not considered in the previous experimental study by McCreesh and
Prosser~\cite{McCreesh:2015}.  This set of instances is also much larger than that of Battiti and
Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm portfolios for subgraph
isomorphism problems.  Battiti and Mascia only considered a pure parallel portfolio consisting of
two solvers, and without a selection mechanism. Their problem set consisted entirely of easy
satisfiable instances.

@@

We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the \Glasgow algorithm~\cite{McCreesh:2015},
the C implementation of \LAD{}~\cite{Solnon:2010}, and the VFLib C
implementation of \VFtwo{}~\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds ($\approx$ 27.8 hours).

@@

\noindent Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
inconsistency, \IncompleteLAD removes candidate pairs. The number of successfully removed pairs
gives information on the distribution of edges (the fewer removed pairs, the more uniform the
distribution). As well as the number of removed pairs, we also record the percentage with respect to
all possible pairs, and the minimum and maximum values on a per-variable basis.

@@

\begin{table}[p]
    \centering\setlength{\tabcolsep}{1em}
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 2219 & 5822809\\
\LLAMA & 705097 & 2203 & 6529563\\
\GlasgowTwo & 1960683 & 2173 & 7783492\\
   \bottomrule \\
\end{tabular}
\caption{Algorithm selection performance on the reduced set of 2336 instances. MCP is the
misclassification penalty; that is, the additional time required to solve an instance because of
choosing solvers that perform worse than the best. Mean performance is over all instances; when an
instance is not solved, its performance is set to the time limit
(\SI{e8}{\ms}).}\label{tab:res}
\end{table}

@@

We train the algorithm selection model specifically for the timeout of $10^8$
milliseconds. In particular, we are interested in minimising the performance
difference to the virtual best. Problem instances that take longer to solver
contribute more to this difference than easy instances and therefore carry more
weight for the algorithm selection model. That is, choosing the wrong solver for
a hard instances is much worse than choosing the wrong solver for an easy
instance.

@@

To gain further insight into the behaviour of the algorithms, we
investigated what affects the relative performance of \PathLAD and \GlasgowTwo.
This pair is of particular interest because they are the best ``medium-case''
algorithms that use strong and weak filtering during search respectively. We
used machine learning techniques to train a simple, human-understandable model
which is able to distinguish these solvers and gives performance better than
always choosing one of them. The model uses four rules:

>>>>

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs~\cite{constraints10}.
\item Classes 2 and 3 contain instances built from a database containing
    various kinds of graph gathered by Larrosa and Valiente~\cite{LV02}: class
    2 contains small instances generated from the first 50 graphs of the
    database, and class 3 contains larger instances with pattern graphs from
    the first 50 graphs of the database and target graphs from the next 50
    graphs.
\item Classes 4 to 8 contain randomly generated graphs from a database of
    graphs commonly used for benchmarking subgraph isomorphism
    algorithms~\cite{GraphDatabase1}: bounded-degree graphs for
    classes 4 and 5, regular meshes for classes 6 and 7, and random graphs with
    uniform edge probabilities for class 8. All of these instances are
    satisfiable.
\item Classes 9 and 10 contain instances from segmented images~\cite{pr15,cviu11}.
\item Class 11 contains instances from meshes modeling 3D objects~\cite{cviu11}.
\item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
    phase transition---these instances are expected to be particularly challenging, despite their
    small size.
\end{itemize}

\noindent Note that Classes 3 and 12 were not considered in the previous
experimental study by McCreesh and Prosser~\cite{McCreesh:2015}.  This set of
instances is also much larger than that of Battiti and
Mascia~\cite{battiti-mascia07}, who were the first to propose algorithm
portfolios for subgraph isomorphism problems.  Battiti and Mascia only
considered a pure parallel portfolio consisting of two randomised solvers, and
without a selection mechanism. Their problem set consisted entirely of
satisfiable instances.

@@

We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the \Glasgow algorithm~\cite{McCreesh:2015},
the C implementation of \LAD{}~\cite{Solnon:2010}, and the VFLib C++
implementation of \VFtwo{}~\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds ($\approx$ 27.8 hours).

@@

\noindent Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
inconsistency, \IncompleteLAD removes candidate pairs. The number of successfully removed pairs
gives information on the distribution of edges (the fewer removed pairs, the more uniform the
distribution). As well as the number of removed pairs, we also record the percentage with respect to
all possible pairs, and the minimum and maximum percentages of removed values on
a per-variable basis.

@@

\begin{table}[p]
    \centering\setlength{\tabcolsep}{1em}
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 2219 & 5822809\\
\LLAMA & 705097 & 2203 & 6529563\\
\GlasgowTwo & 1960683 & 2173 & 7783492\\
   \bottomrule \\
\end{tabular}
\caption{Algorithm selection performance on the reduced set of 2336 instances. MCP is the
misclassification penalty; that is, the additional time required to solve an instance because of
choosing solvers that perform worse than the best. Mean MCP and performance are over all 2336 instances; when an
instance is not solved, its performance is set to the time limit
(\SI{e8}{\ms}).}\label{tab:res}
\end{table}

@@

We train the algorithm selection model specifically for the timeout of $10^8$
milliseconds. In particular, we are interested in minimising the performance
difference to the virtual best. Problem instances that take longer to solve
contribute more to this difference than easy instances and therefore carry more
weight for the algorithm selection model. That is, choosing the wrong solver for
a hard instances is much worse than choosing the wrong solver for an easy
instance.

@@

To gain further insight into the behaviour of the algorithms, we
investigated what affects the relative performance of \PathLAD and \GlasgowTwo.
This pair is of particular interest because they are the best ``medium-case''
algorithms that use strong and weak filtering during search respectively. We
used machine learning techniques to train a simple, human-understandable model
which is able to distinguish these solvers for the 2336 hard instances and gives
performance better than always choosing one of them. The model uses four rules:
